---
title: "Tidy Tuesday - Incarceration Rate"
author: "David Izydorczyk"
date: "`r format(Sys.time(), "%a %b %d ")`"
output: github_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, I start by loading the tidyverse library & the data. 

```{r}
library(tidyverse)
library(magrittr)
library(gganimate)

prison <- read.csv("../data/prison_population.csv")


```

Next, I am going to convert the data.frame into a tibble and inspect the data. 

```{r}
(prison %<>% as_tibble())
```
Plot the 15 states with the highest incarceration rates

```{r}


prison %>% 
  mutate(lumped = fct_lump(state,25)) %>%
  count(lumped, sort = TRUE) %>%
  filter(lumped != "Other") %>%
  mutate(state_lumped = fct_reorder(lumped,n)) %>%
  ggplot(.,aes(x = state_lumped,y = n )) + 
    geom_col()+
    coord_flip() +
    labs(x = "N",
         y = "State")

```

We see that texas has the highest incarceration rate over of all states. 

Lets have a look at the incarceration rate by urbanicity.

```{r}
prison %>% 
  group_by(urbanicity) %>%
  count()
```

Next, I want to see how the prison population changed of the years.

```{r}
theme_set(theme_light(),
          base_size = 18)

prison %>% 
  filter(pop_category == "Total") %>%
  group_by(year) %>%
  summarize(prison_pop_m = mean(prison_population,na.rm=TRUE),
            prison_pop_sd = sd(prison_population,na.rm=TRUE)) %>% 
  ggplot(., aes(x = year, y = prison_pop_m)) +
  geom_line() +
  labs(x = "Year",
       y = "Mean Prison Population")
```

Next we try a moving bubble charta ala "Factfullness" to see the development of prison population for different types. 


```{r}
theme_set(theme_light(base_size = 18))






a <- prison %>% 
  filter(year>=1990) %>%
  group_by(year,pop_category) %>%
  filter(pop_category == "Native American" |
           pop_category == "Asian" |
           pop_category == "Black" |
           pop_category == "Latino"|
           pop_category == "White") %>% 
  summarize(prison_pop_m = mean(prison_population,na.rm=TRUE),
            pop_size     = mean(population,na.rm=TRUE)) %>% 
  filter(!is.nan(prison_pop_m)) %>%
  ggplot(.,aes(x = year, y = prison_pop_m, color = pop_category, group = pop_category)) +
    geom_line() + 
    geom_segment(aes(xend = 2015, yend = prison_pop_m), linetype = 2, colour = 'grey') + 
    geom_point(aes(x = year,size = pop_size)) + 
    geom_text(aes(x = 2016, label = pop_category), hjust = 0) +
    labs(x = 'Year',
         y = 'Prison Population Size ',
         title = "Development of the prison population size for different populations",
         color = "Population Category",
         size = "Population Size") +
    transition_reveal(year) +
    coord_cartesian(clip = 'off') 



animate(a, height = 400, width =700,nframes=100)
anim_save(filename ="../figures/tt_incarceration2.gif")
```


```{r}
theme_set(theme_light(base_size = 17))

library(tidyverse)
library(magrittr)
library(gganimate)

prison <- read.csv("../data/prison_population.csv")

states <- prison %>% 
            group_by(state) %>%
            summarize(prison_pop_m = mean(prison_population,na.rm=TRUE),
                      pop_size     = mean(population,na.rm=TRUE)) %>%
             top_n(n=5,wt = prison_pop_m) %>%
             .$state  %>% as.character()


a <- prison %>% 
  filter(year>=1990) %>%
  filter(state %in% states) %>% 
   group_by(year,state) %>%
  summarize(prison_pop_m = mean(prison_population,na.rm=TRUE),
            pop_size     = mean(population,na.rm=TRUE)) %>% 
  filter(!is.nan(prison_pop_m)) %>%
  ggplot(.,aes(x = year, y = prison_pop_m, color = state, group = state)) +
    geom_line() + 
    geom_segment(aes(xend = 2015, yend = prison_pop_m), linetype = 2, colour = 'grey') + 
    geom_point(aes(x = year,size = pop_size)) + 
    geom_text(aes(x = 2016, label = state), hjust = 0) +
    xlim(1990,2018) +
    labs(x = 'Year',
         y = 'Prison Population Size ',
         title = "Development of the prison population size",
         subtitle = " In the five states with the highest prison population size",
         color = "State",
         size = "Population Size") +
    transition_reveal(year) +
    coord_cartesian(clip = 'off') 



animate(a, height = 400, width =700,nframes=100)
anim_save(filename ="../figures/tt_incarceration2.gif")
```
